---
title: "Introduction to Monte Carlo Simulation"
author: "Tyler J. Brough, Ph.D."
date: "October 11, 2017"
output: html_document
---

## A Simple Example with the Exponential Distribution

```{R}
# Set the sample size 
N <- 200
lam = 4.5

# Number of repititions for the Monte Carlo study
M <- 200000

# The Main Monte Carlo loop
avgs <- rep(0, M)
for(i in 1:M)
{
  x = rexp(N, rate = lam)
  avgs[i] = mean(x)
}

# Make a density plot
plot(density(avgs), lwd = 2, type = "l", col = "blue")
```


```{R}
xbar <- mean(avgs)
s <- sd(avgs)
```


## Simulating the Simple Linear Regression Model

The Simple CLRM is the following:

$$
y_{i} = \alpha + \beta x_{i} + \varepsilon_{i}
$$

We can simulate the CLRM by choosing some parameter values and simulating values of $X$ and $\varepsilon$

```{R}
alpha <- 2.0
beta <- 3.2
x <- runif(M)
u <- rnorm(M)
y <- alpha + beta * x + u
head(y, 20)
```

Let's simulate the sampling distribution for the CLRM:


```{R}
R <- 500
bhat <- rep(0, R)

for(i in 1:R)
{
  u <- rnorm(N)
  y <- alpha + beta * x + u
  fit <- lm(y ~ x)
  bhat[i] <- coef(fit)[[2]]
}

plot(density(bhat), lwd = 3, type = "l", col = "green")
```



## The Multiple CLRM

The multiple CLRM is the following:

$$
y_{i} = \alpha + \beta_{1}x_{1,i} + \beta_{2} x_{2,i} + \epsilon_{i}
$$

Let's simulate this process as follows (choosing some arbitrary parameters). First let's examine the concept of ommitted variable bias

```{R}
R <- 100
alpha <- 1.7
beta1 <- 2.8
beta2 <- 3.2
x1 <- rnorm(N, sd = 20)
x2 <- rnorm(N, sd = 0.8)

b1hat <- rep(0, R)
for(i in 1:R)
{
  # the "true" data generating process
  u <- rnorm(N)
  y = alpha + beta1 * x1 + beta2 * x2 + u
  
  # the econometricians estimate (omitting x2)
  b1hat[i] <- coef(lm(y ~ x1))[[2]]
}

b1avg <- mean(b1hat)
b1sd <- sd(b1hat)
print(b1avg, b1sd)


## Make a density plot
plot(density(b1hat), lwd = 3, type = "l", col = "red")
```


## HWK Problem 2

```{r}
alpha = 0.0
beta = 0.97
```

Then you can run a single run of the simulation as follows:

```{r}
x <- runif(100)
u <- rnorm(100)
y <- alpha + beta * x + u
```


```{r}
N <- 1000





# Set up the simulation
x <- runif(N)
#bhat <- rep(0, N)
u <- rnorm(N)
y <- alpha + beta * x + u
  
# Hand data to econometrician
fit <- lm(y ~ x)
obj <- summary(fit)


# The econometrician forms the hypothesis test
bhat <- coef(fit)[2]
std.err <- obj$coefficients[,2][2]
tstat <- (bhat - 1.0) / std.err
tcrit <- 2.08
```

